---
title: "Prediction Assignment Writeup"
author: "N Vinod Kumar"
date: "15/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Loading
```{r}
library(caret)
library(randomForest)
setwd("\\Coursera\\MachineLearning\\Project")
training <- read.csv("pml-training.csv",header = TRUE)
testing <- read.csv("pml-testing.csv",header = TRUE)
```

### Selecting only the Required Columns
The goal for the prediction is use to the data from accelerometers on the belt, forearm, arm, and dumbbell. We select only these columns from the training data,
```{r}
training <- training[8:160]
```

### Removing the columns where all are NA
```{r}
training <- training[,which(unlist(lapply(training, function(x) all(!is.na(x)))))]
```

### Removing the columns where all are #DIV/0!
```{r}
training <- training[,which(unlist(lapply(training, function(x) all(!x=="#DIV/0!"))))]
```

### Removing highly correlated columns
```{r}
cordata <- cor(training[c(1:52)] ,use="complete.obs")
findCorrelation(cordata, cutoff = 0.90, verbose = TRUE, names=TRUE)
```


```{r}
cols <- findCorrelation(cordata, cutoff = 0.90, verbose = TRUE, names=TRUE) 
training <- training[,!(names(training) %in% cols)]
```

### Building Model with Regression tree
```{r}
model1 <- train(classe ~ ., method="rpart", data=training)
model1
```

## Model Prediction
```{r}
pred1 <- predict(model1, testing)
pred1
```

### Building Model with Random Forests
```{r}
# model2 <- train(classe ~ ., method="ranger", data=train)
```
Growing trees.. Progress: 93%. Estimated remaining time: 2 seconds.
Growing trees.. Progress: 80%. Estimated remaining time: 7 seconds.
Growing trees.. Progress: 94%. Estimated remaining time: 1 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 99%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 93%. Estimated remaining time: 2 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.


### Testing the model
```{r}
# pred2 <- predict(model2, testing)
# pred2
```
[1] B A B A A E D B A A B C B A E E A B B B
Levels: A B C D E