---
title: "Prediction Assignment Writeup"
author: "N Vinod Kumar"
date: "15/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Loading
```{r}
library(caret)
library(randomForest)
setwd("\\Coursera\\MachineLearning\\Project")
training <- read.csv("pml-training.csv",header = TRUE)
testing <- read.csv("pml-testing.csv",header = TRUE)
```

### Selecting only the Required Columns
The goal for the prediction is use to the data from accelerometers on the belt, forearm, arm, and dumbbell. We select only these columns from the training data,
```{r}
training <- training[8:160]
```

### Removing the columns where all are NA
```{r}
training <- training[,which(unlist(lapply(training, function(x) all(!is.na(x)))))]
```

### Removing the columns where all are #DIV/0!
```{r}
training <- training[,which(unlist(lapply(training, function(x) all(!x=="#DIV/0!"))))]
```

### Removing highly correlated columns
```{r}
cordata <- cor(training[c(1:52)] ,use="complete.obs")
findCorrelation(cordata, cutoff = 0.90, verbose = TRUE, names=TRUE)
```


```{r}
cols <- findCorrelation(cordata, cutoff = 0.90, verbose = TRUE, names=TRUE) 
training <- training[,!(names(training) %in% cols)]
```

### Building Model with Regression tree
```{r}
model1 <- train(classe ~ ., method="rpart", data=training)
model1
```

## Model Prediction
```{r}
pred1 <- predict(model1, testing)
pred1
```

### Building Model with Random Forests
```{r}
# model2 <- train(classe ~ ., method="ranger", data=train)
```
Growing trees.. Progress: 93%. Estimated remaining time: 2 seconds.
Growing trees.. Progress: 80%. Estimated remaining time: 7 seconds.
Growing trees.. Progress: 94%. Estimated remaining time: 1 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 99%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 93%. Estimated remaining time: 2 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 100%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 0 seconds.


### Testing the model
```{r}
# pred2 <- predict(model2, testing)
# pred2
```
[1] B A B A A E D B A A B C B A E E A B B B
Levels: A B C D E

### Confusion Matrix
```{r}
#confusionMatrix(pred2, TestingSet$classe)
```

Confusion Matrix and Statistics
           Reference
 Prediction    A    B    C    D    E
          A 1674    7    0    0    0
          B    0 1131    6    0    0
          C    0    1 1020    5    0
          D    0    0    0  958    1
          E    0    0    0    1 1081
 
 Overall Statistics
                                           
                Accuracy : 0.9964          
                  95% CI : (0.9946, 0.9978)
     No Information Rate : 0.2845          
     P-Value [Acc > NIR] : < 2.2e-16       
                                           
                   Kappa : 0.9955          
  Mcnemar's Test P-Value : NA              
 
 Statistics by Class:
 
                      Class: A Class: B Class: C Class: D Class: E
 Sensitivity            1.0000   0.9930   0.9942   0.9938   0.9991
 Specificity            0.9983   0.9987   0.9988   0.9998   0.9998
 Pos Pred Value         0.9958   0.9947   0.9942   0.9990   0.9991
 Neg Pred Value         1.0000   0.9983   0.9988   0.9988   0.9998
 Prevalence             0.2845   0.1935   0.1743   0.1638   0.1839
 Detection Rate         0.2845   0.1922   0.1733   0.1628   0.1837
 Detection Prevalence   0.2856   0.1932   0.1743   0.1630   0.1839

# Results
The accuracy of Random Forest is 99.64, which is higher than that by Generalized Boosted. The Random Forest
 Balanced Accuracy      0.9992   0.9959   0.9965   0.9968   0.9994